---
title: "results_summary"
author: "Miguel Camacho"
date: "20/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(phyloseq)
library(tidyverse)
library(metagMisc)
library(here)
```

```{r wd, echo=T}
here::i_am("code/generate_outputs/results_summary.Rmd")
```

## sequencing results

## alpha diversity

```{r alpha diversty, echo=T}
ps <-
  readRDS(file.path(here(),"data/intermediate/ps_t_noRep.rds"))
asv_per_sample <- otu_table(ps) %>% as.data.frame() %>% apply(1, function(x) sum(x > 0))
```

A final `r phyloseq::ntaxa(ps)` ASVs were determined, ranging from `r min(asv_per_sample)` to `r max(asv_per_sample)` (sd = `r sd(asv_per_sample)`).

## coverage per taxonomic rank

```{r tax coverage, echo=T}
source(file.path(here(),"code/functions/coverage_rank.r"))
rcov <- rank_cov(ps)
rcov
```

##  models alpha diversity
### effect of season

```{r model_alpha_season}

```
# variance across seasons

We wanted to check the variance in Shannon diversity across seasons. For that reason I computed Levene`s test for the inequality of variance using nested factors: that is, the four values for each type of soil in each species plot.

```{r levene test}
load(file.path(here(),"data/intermediate/levene.Rdata"))
plev <- levtest$`Pr(>F)`[1]
flev <- levtest$`F value`[1]
```

Variance of alpha diversity was unequal across seasons depending on the species (F9 = `r round(flev, 2)`, p = `r round(plev, 2)`), with a more stable diversity in the plots of Spartina maritima (sdSpartina_maritma = `r filter(d1, species == "Spartina_maritima") %>% pull(sd) %>% mean()`; sdother plots = `r filter(d1, species != "Spartina_maritima") %>% pull(sd) %>% range()`).

# Archaea and Bacterial diversity
```{r}
ps_t <- transform_sample_counts(ps, function(x) x / sum(x))
ps_bact <- phyloseq::subset_taxa(ps_t, domain %in% "Bacteria")
ps_arch <- phyloseq::subset_taxa(ps_t, domain %in% "Archaea")
ps_domain <- tax_glom(ps_t, "domain")
```

Of the `r ntaxa(ps)`, `r ntaxa(ps_arch)` ASVs (`r otu_table(ps_domain)[,1] %>% { . * 100} %>% mean() %>% round(2)` % of overall relative abundance) corresponded to Archaea and `r ntaxa(ps_bact)` to Bacteria. All of the Archaea corresponded to halophylic taxa from multiple genera within the Order Halobactorales, Phylum Holobacterota.